---
title: "Alcohol vs Seat choice"
author: "Ana Mendez"
date: "Spring 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
require(tidyverse)
require(tigerstats)
davisdata <- read.csv(file="UCDavis.csv")
```
## Our Goal is to answer this question Is there a difference between the alcohol consumed between the row
## you choose to sit in??

## Now with ggplot - first select the basic data

```{r}
basicCN <- ggplot(davisdata,aes(y=alcohol,x=Seat))
```
## Now add in boxplot

```{r}
basicCN + geom_boxplot()
```
  
  The medium number of drinks is higher than the people seated at the front and the people seated in the middle row. We can see that by looking at the boxes. theres not too much difference between the back and the middle and front 
  
### get fancy  

```{r}
basicCN + geom_boxplot(notch=TRUE,outlier.color = "red") + stat_boxplot(geom ='errorbar') + labs(title="Fancy notch boxplot \n with 2 lines")
```

The fancy notch displays the three categories of seating arrangements which was the front, middle and back. We can observe that people who are seated in the back had a median that differed while the people in the front and back had more confident median plot.
We can assume that people in the back consume more alcoholic drinks compared to those who were seated in the middle and front rows.

## Numerical Summarys 

```{r}
favstats(alcohol~Seat,data=ucdavis1)
```

The numerical data shows that the average number of drinks for the people seated in the back is about 9 drinks.
This is three times the average number of drinks for those who are seated in the middle. 
This is also nine times average number of drinks for those who are seated in the front.


## Inferential Tests

```{r}
model1 <- aov(alcohol~Seat,data=davisdata)
summary.aov(model1)
```

The one way anova inferential test  shows a small p-value which is 0.000116, which means that there is a probability of having no difference between alcohol consumed between the front, middle and back rows that are available for people to be seated in. We reject the null hypothesis becuase the probability of getting this result is significantly lower

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

This plot shows that there is a difference in the alcohol consumed between the three sections that a person can be seated in.
The intervals from the Middle-Front cross which tells us that there is not a significant difference between these two sections. on the other hand, the intervals of the Front-Back and Middle-Back never cross which means that there is a significant difference between the amount of drinks per section.


## Conclusions

 After looking over the numerical data and inferential tests, we can conclude that the people who are seated in the back section are consuming more alcohol that those who are seated in the front or middle section. The numerical data shows that the average number of drinks for the people seated in the back is about 9 drinks while there is three times the average number of drinks for those who are seated in the middle and nine times average number of drinks for those who are seated in the front. We also got a p-value of 0.000116 which indicates that there is a chance of having no difference between alcohol consumed between the front, middle and back rows that are available for people to be seated in.

